---
title: "Thermal comfort analysis of Cape Town, South Africa"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(terra)
library(tidyterra)
library(rgee)
library(sf)
library(here)
```

This notebook documents the thermal comfort analysis for an area of interest in downtown Cape Town.

# Load datasets

MRT, shadow, and boundary data were downloaded from the [wri-cities-heat bucket in AWS](https://us-east-1.console.aws.amazon.com/s3/buckets/wri-cities-heat?prefix=ZAF-Cape_town/&region=us-east-1&bucketType=general). LULC was downloaded from [GEE](https://code.earthengine.google.com/?asset=projects/wri-datalab/cities/SSC/LULC_V2). Mean summertime (Dec, Jan, Feb) LST was calculated and downloaded from [GEE](https://code.earthengine.google.com/61495c5b32e5f4a86f2e5e3114618f15).

```{r}
mrt <- rast(here("data", 
                 "ZAF-Cape_town", 
                 "processed", 
                 "solweig", 
                 "Tmrt_2022_22_1400D.tif")) %>% 
  rename(MRT = Tmrt_2022_22_1400D) 

shadow <- rast(here("data", 
                    "ZAF-Cape_town", 
                    "processed", 
                    "solweig", 
                    "Shadow_2022_22_1400D.tif")) %>% 
  rename(Shadow = Shadow_2022_22_1400D) 
shadow <- shadow * 10
shadow <- shadow %>% 
  as.factor()

boundary <- st_read(here("data", 
                         "ZAF-Cape_town", 
                         "processed", 
                         "area_of_interest.geojson"))

lulc <- rast(here("data", 
                  "ZAF-Cape_town", 
                  "processed", 
                  "ZAF-Cape_town-LULCV2.tif")) %>% 
  as.factor()

lulc.solweig <- rast(here("data", 
                  "ZAF-Cape_town", 
                  "processed", 
                  "landcover_nozero.tif")) %>% 
  as.factor()

lst <- rast(here("data", 
                 "ZAF-Cape_town", 
                 "processed", 
                 "ZAF-Cape_town-LST-summer2021-22.tif"))
```

# Mean radiant temperature

```{r}
ggplot() +
  geom_sf(data = boundary, fill = NA) +
  geom_spatraster(data = mrt) +
  scale_fill_whitebox_c(
    palette = "soft",
    labels = scales::label_number(suffix = "ยบ"),
    n.breaks = 8,
    guide = guide_legend(reverse = TRUE, title = "MRT")) +
  ggtitle("Mean radiant temperature, Cape Town City Centre")

mrt %>% 
  as_tibble() %>% 
ggplot() +
  geom_density(aes(MRT)) +
  labs(title = "Distribution of MRT, 24 January 2022",
       subtitle = "Cape Town City Centre")
```

# Land surface temperature

```{r}
ggplot() +
  geom_sf(data = boundary, fill = NA) +
  geom_spatraster(data = lst) +
  scale_fill_whitebox_c(
    palette = "soft",
    labels = scales::label_number(suffix = "ยบ"),
    n.breaks = 8,
    guide = guide_legend(reverse = TRUE, title = "LST")) +
  labs(title = "Mean summertime land surface temperature, Cape Town City Centre",
       subtitle = "Dec 2021, Jan 2022, Feb 2022")

lst %>% 
  as_tibble() %>% 
ggplot() +
  geom_density(aes(LST)) +
  labs(title = "Density of summertime (2021-22) LST",
       subtitle = "Cape Town City Centre")
```
# Shade

```{r}
ggplot() +
  geom_sf(data = boundary, fill = NA) +
  geom_spatraster(data = shadow) +
  scale_fill_manual(values = c("#765341", "#0B6623", "#FFFFFF"),
                    labels = c("Building", "Tree", "None"),
                    name = "Shade",
                    na.value = "transparent") +
  labs(title = "Shade, Cape Town City Centre",
       subtitle = "14:00 local time")

shadow %>% 
  as_tibble() %>% 
ggplot() +
  geom_bar(aes(Shadow, fill = Shadow)) +
  scale_fill_manual(values = c("#765341", "#0B6623", "#FFFFFF"),
                    labels = c("Building", "Tree", "None"),
                    name = "Shade",
                    na.value = "transparent") +
  labs(title = "Shaded area at 14:00 local time",
       subtitle = "Cape Town City Centre") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

# MRT by LULC

```{r}
mrt.lulc <- zonal(mrt, lulc.solweig)
mrt.lulc
```

